---
title: "Optimal Designs for Detecting Hormesis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Optimal Designs for Detecting Hormesis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nlodm)
```

## Introduction

This vignette recreates the optimal designs found in Dette, Pepelyshev, Wong (2011) and the draft under preparation.

## Models

The Hunt-Bowman model is defined as $$
\mu(d, \theta) = \begin{cases} c_1d^2 - c_1\tau d + \frac{1}{1+e^{\beta_0}}& 0 \leq d \leq \tau\\
\frac{1}{1 + e^{\beta_0 - \beta_1(d-\tau)}} & \tau > d\end{cases} 
$$ The parameter vector is $\theta = (c_1, \tau, \beta_0, \beta_1)^T$.

The gradient is $$
\nabla \mu(d, \theta) = \begin{cases} (d^2 - \tau d, -c_1d, - \frac{e^{\beta_0}}{(1+e^{\beta_0})^2}, 0)^T & 0 \leq d \leq \tau\\
\left(0, \beta_1 \frac{e^{\beta_0-\beta_1(d-\tau)}}{(1+e^{\beta_0-\beta_1(d-\tau)})^2}, -\frac{e^{\beta_0-\beta_1(d-\tau)}}{(1+e^{\beta_0-\beta_1(d-\tau)})^2}, \frac{(d-\tau)e^{\beta_0-\beta_1(d-\tau)}}{(1+e^{\beta_0-\beta_1(d-\tau)})^2} \right)^T & \tau > d
\end{cases}
$$

Define a custom gradient function:

```{r}
grad.huntbowman = function(x, theta) {
  
  c1 = theta[1]
  tau = theta[2]
  b0 = theta[3]
  b1 = theta[4]
  
  if (x <= tau) {
    return(c(
      x^2 - tau*x,
      -c1*x,
      -exp(b0)/(1+exp(b0))^2,
      0
    ))
  }
  else if (x > tau) {
    
    f = exp(b0 - b1*(x - tau))/(1 + exp(b0-b1*(x-tau)))^2
    return(c(
      0,
      -b1*f,
      -f,
      (x - tau) * f
    ))
  }
}
```

The exp-log model is defined as $$
\mu(d, \theta) = c_0e^{-c_1d} + \frac{1}{1+e^{\beta_0 - \beta_1d}}
$$

and the parameter vector is $\theta = (c_0, c_1, \beta_0, \beta_1)^T$.

The gradient is

$$
\nabla \mu(d, \theta) = \left(e^{-c_1 d}, -c_0 d e^{-c_1d}, -\frac{e^{\beta_0-\beta_1d}}{(1+e^{\beta_0-\beta_1d})^2}, \frac{de^{\beta_0-\beta_1d}}{(1+e^{\beta_0-\beta_1d})^2}\right)^T
$$

Custom gradient function:

```{r}
grad.logexp = function(x, theta) {
  
  c0 = theta[1]
  c1 = theta[2]
  b0 = theta[3]
  b1 = theta[4]
  
  f = exp(b0-b1*x)/(1+exp(b0-b1*x))^2
  
  return(c(
    exp(-c1*x),
    -c0*x*exp(-c1*x),
    -f,
    x*f
  ))
}
```

## Designs for Hunt-Bowman model

Approximate locally D-optimal design:

```{r}
hb_D = nlodm(
  grad_fun = grad.huntbowman,
  obj = 'D',
  theta = c(170, 0.04, 1.46, 40),
  bound = 0.15,
  pts = 4,
  algorithm = 'DE',
  swarm = 50,
  iter = 500,
  seed = 1234,
)
```

```{r}
hb_D$plot
```

```{r}
hb_D$design
```

Exact locally D-optimal design:

```{r}
hb_D_exact = nlodm(
  grad_fun = grad.huntbowman,
  obj = 'D',
  theta = c(170, 0.04, 1.46, 40),
  bound = 0.15,
  pts = 20,
  algorithm = 'PSO',
  swarm = 100,
  iter = 1000,
  seed = 1234,
  exact = T
)
```

```{r}
hb_D_exact$design
```

## Exp-log model
