---
title: "Finding Multi-objective Designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multi-objective}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nlodm)
```

## Introduction

It is often the case that more than one objective must be considered when designing an experiment. For example, when designing an experiment to estimate the benchmark dose (BMD) in toxicology, it is helpful to consider the trade-offs between estimating all of the model parameters and the BMD. In Gertsch and Wong (2023), this was done by combining c and D objectives into a compound objective. Another example comes from [Sanchez, Sarabia, and Ortiz (2012)](https://doi.org/10.1016/j.aca.2012.10.014) where the variance inflation factors of 7 parameters are considered.

There are two major approaches to multi-objective design:

1.  Compound design criteria

2.  Pareto front analysis

We will briefly introduce both of these approaches and give examples of how to use the `nlodm` package to find these designs.

## Compound optimal designs

The basic idea behind compound design criteria is to construct a criterion from a convex combination of convex objective functions. Let $\Psi_i$, $i = 1, \dots, k$ be $k$ convex design criteria and let $\alpha_i$ be weights such that $\sum_i^k \alpha_i = 1$. A compound criteria can then be defined as

$$
\Psi(M) = \sum_i^k \alpha_i \Psi_i(M)
$$

Similar to (pseudo) Bayesian criteria, there is an equivalence theorem that applies these compound criteria, meaning that is possible to check if the design found is optimal.

### BMD designs

In the single objective vignette for this package, a design criteria was derived for estimating the benchmark dose. This criteria was a combination of D and c-optimality.

$$
\Psi_{BMD}^\lambda(M) = \frac{(1-\lambda)}{p} \log (|M|) - \lambda \log \left(c' M^{-1} c \right)
$$

This specific type of dual objective design is implemented in the `nlodm` function.

```{r, results='hide'}
out_bmd = nlodm(
  model = 'Log-logistic',
  grad_fun = grad.loglogistic,
  obj = 'bmd',
  theta = c(0.02461,-2.390, 1),
  bound = 30,
  pts = 3,
  algorithm = 'DE',
  swarm = 50,
  iter = 500,
  seed = 1234,
  bmd_type = 'added',
  risk = 0.1,
  lambda = 0.5
)
```

```{r}
out_bmd$design
```

```{r}
out_bmd$plot
```

### Custom compound designs

A major goal of this package is allow the user to define custom compound design criteria for any number of objectives. We can do this using the `multi_obj` function. As an example, we will implement BMD designs using this function.

## Pareto optimal designs

The major downside of compound design objectives is that the weighting parameter(s) must be chosen prior to solving the optimization problem. Practically, this means assigning priorities to the different objectives before optimization, which is difficult because it may not be obvious how different values of the weight parameter(s) correspond to the desired priorities. For example, in a dual objective problem with a single weight parameter $\lambda$, setting $\lambda=0.5$ does not necessarily imply that each objective will be given equal priority. This can be seen in the efficiency plot in the previous section (to be finished).

An alternative approach that is popular in engineering is to construct a Pareto front of possible solutions to the problem. Intuitively, the Pareto front represents the optimal trade-offs between the multiple objectives. This allows the decision about objective importance to be made post-hoc.

### Introduction to Pareto fronts

TODO

### Example: designing for BMD estimation using the Pareto front

In previous examples, we looked at dual D and c-optimal designs using a compound criteria. We will be using the $c_\epsilon$ criterion to help avoid matrix singularities. Now let's use a true multi-objective optimization approach and generate a Pareto front for these 2 objectives.

Set up:

```{r}
grad_funs = list(grad.loglogistic, grad.loglogistic)
obj_funs = list(obj.D, obj.c_e)
bmd_grad = get_bmd_grad("Log-logistic", 'added')
theta = c(0.02461, -2.390, 1)
thetas = list(
    theta,
    theta
  )
params = list(
    c(),
    bmd_grad(0.1, theta)
  )
```

Call optimizer:

```{r}
result = multi_obj(
    grad_funs = grad_funs,
    obj_funs = obj_funs,
    thetas = thetas,
    params = params,
    type = 'pareto',
    bound = 30,
    pts = 3,
    swarm = 100,
    maxiter = 500,
    verbose = F
  )
```

```{r}
pareto_data = extract_front(result)
plot_pareto2d(pareto_data, c("D", "c_e"))
```

It is helpful to find the c and D-optimal designs and add them to the plot. The c-optimal design is likely singular, so we will try to find the $c_\epsilon$ design. We will also add the BMD design found earlier using the compound criterion.

```{r}
out_D = nlodm(
  model = NULL,
  grad_fun = grad.loglogistic,
  obj = 'D',
  theta = theta,
  bound = 30,
  pts = 3,
  algorithm = 'PSO',
  swarm = 30,
  iter = 400,
  seed = 939
)
```

```{r}
out_ce = nlodm(
  model = NULL,
  grad_fun = grad.loglogistic,
  obj = 'c_e',
  theta = theta,
  bound = 30,
  pts = 3,
  algorithm = 'PSO',
  swarm = 30,
  iter = 400,
  seed = 939,
  c = bmd_grad(0.1, theta)
)
```

```{r}
D_obj = obj_fun_factory(grad.loglogistic, obj.D, matrix(theta,nrow=1), param=c())
ce_obj = obj_fun_factory(grad.loglogistic, obj.c_e, matrix(theta,nrow=1), 
                         param=matrix(bmd_grad(0.1, theta), nrow = 1))

plot_pareto2d(pareto_data, c("D", "c_e")) +
  annotate('point', x = -D_obj(out_D$raw$result), y = -ce_obj(out_D$raw$result), color = 'red') +
  annotate('point', x = -D_obj(out_ce$raw$result), y = -ce_obj(out_ce$raw$result), color = 'red') +
  annotate('point', -D_obj(out_bmd$raw$result), y = -ce_obj(out_bmd$raw$result), color = 'purple')
```

### Example: 3 objectives

### Example: 4 objectives
